Constraints on Assembly Bias from Galaxy Clustering.

### 2016-06-08 ms.tex
Rebased commit on master to resolve conflicts

\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\vmax}{v_{\rm max}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\ncen}{N_{\rm cen}}
\newcommand{\nsat}{N_{\rm sat}}
\newcommand{\ngal}{N_{\rm gal}}
\newcommand{\abias}{\mathcal{A}_{\rm bias}}

@@

The ultimate goal of halo occupation modeling is to characterize the
galaxy-dark matter connection by constraining the probability
distribution $P(X_{\rm g}|Y_{\rm h})$. Here $X_{\rm g} = {x_1, x_2,
  ...x_n}$ is a set of $n$ galaxy properties, and $Y_{\rm h} = {y_1,
  y_2,...,y_m}$ a set of $m$ halo properties. Accurate knowledge of
this probability distribution yields invaluable insight regarding the
processes of galaxy formation, and allows one to make accurate
predictions for the clustering of galaxies as function of any property
$x_i$. Most halo occupation studies to date have restricted themselves
to only one halo property, namely halo mass.  This includes Halo
Occupation Distribution (HOD) models, which aim to constrain
$P(N|\mvir)$, the probability that a halo of mass $M$ contains $N$
galaxies of some specified properties
\citep[e.g.,][]{kauffmann_etal97,jing98,benson01,berlind02}, and
Conditional Luminosity Function (CLF) models, which aim to constrain
$\Phi(L|M)$, the luminosity (or stellar mass) function of galaxies in
haloes of mass $M$ \citep[e.g.,][]{yang03,vdBosch03a,cooray06}. The
HOD and CLF models are mainly used in combination with the halo model,
which describes the non-linear matter field in terms of halo building
blocks \citep[e.g.,][]{seljak00,cooray02}, to predict the clustering
of galaxies as function of luminosity or stellar mass.  Although
entirely analytical, and therefore fast, the downside of this method
is that it is extremely challenging to construct a model that is
accurate at the few percent level. The main issues are halo exclusion
\citep[see][]{smith_etal07,vdBosch13}, scale dependence of the halo
bias \citep[see][]{tinker05}, and the halo assembly bias mentioned
above.

These issues are trivially avoided by directly populating dark matter
haloes in numerical simulations with galaxies. This can be done using
either a parameterized model for the halo occupation statistics
\citep[e.g.,][]{Yang_etal04,moster10;behroozi10,behroozi13} or by
rank-order matching galaxy properties (i.e., luminosity or stellar
mass) to halo properties (i.e., halo mass). The latter method is known
as subhalo abundance matching (SHAM).  Although many SHAM studies use
halo mass as the halo property of choice
\citep[e.g.,][]{vale_ostriker04,vale_ostriker06,conroy07,conroy_wechsler09},
a number of studies have used alternative halo properties, such as the
maximum circular velocity, $\vmax$, or its peak value over the halo's
history, $V_{\rm peak}$
\citep[e.g.][]{tasitsiomi_etal04,conroy06,marin_etal08,reddick_etal13}

Since, at fixed halo mass, $\vmax$ is correlated with halo
concentration, and thus with halo assembly time, the resulting
occupation models have assembly bias build in
\citep[see][]{Zentner_etal14}; i.e., at fixed halo mass, galaxy
properties will be correlated with halo assembly
time. \citet{lehmann_etal15} took this one step further, and explored
SHAM models in which the halo property varies between $M$ and $\vmax$
via a continuously-valued parameter. Using SDSS clustering
measurements to constrain their model, they conclude that galaxy
luminosity reveals a significant dependence on halo
concentration. Since the latter is strongly correlated with halo
formation time \citep[e.g.,][]{wechsler02,zhao_etal09,ludlow_etal13},
their conclusions argue for non-zero assembly bias in the galaxy
population.

@@

Our work is important for several reasons. Our work is a re-analysis
of the SDSS DR7 data that overcomes shortcomings of previous
analyses. In particular, we use direct population of galaxies in a
cosmological simulation so that delicate issues present in analytic
modeling, such as scale-dependent halo bias, are treated exactly.  The
simulation we use is based on the latest Planck cosmological
parameters, updating previous work. Furthermore, many important
differences between this work and the previous analysis by
\citet{zehavi_etal11} result from the fact that the Monte Carlo Markov
Chains used by Zehavi \etal where too small and had therefore not
converged (Z. Zheng \& I. Zehavi, Private Communication).  Therefore,
our analysis {\em supersedes previous analyses even in the case of
  standard halo occupation models}.

>>>>

\newcommand{\mvir}{M_{\rm vir}}
\newcommand{\vmax}{V_{\rm max}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\ncen}{N_{\rm cen}}
\newcommand{\nsat}{N_{\rm sat}}
\newcommand{\ngal}{N_{\rm gal}}
\newcommand{\abias}{\mathcal{A}_{\rm bias}}

@@

The ultimate goal of halo occupation modeling is to characterize the
galaxy-dark matter connection by constraining the probability
distribution $P(X_{\rm g}|Y_{\rm h})$. Here $X_{\rm g} = \{x_1, x_2,
  ...x_n\}$ is a set of $n$ galaxy properties, and $Y_{\rm h} = \{y_1,
  y_2,...,y_m\}$ a set of $m$ halo properties. Accurate knowledge of
this probability distribution yields invaluable insight regarding the
processes of galaxy formation, and allows one to make accurate
predictions for the clustering of galaxies as function of any property
$x_i$. Most halo occupation studies to date have restricted themselves
to only one halo property, namely halo mass.  This includes Halo
Occupation Distribution (HOD) models, which aim to constrain
$P(N|\mvir)$, the probability that a halo of mass $M$ contains $N$
galaxies of some specified properties
\citep[e.g.,][]{kauffmann_etal97,jing98,benson01,berlind02}, and
Conditional Luminosity Function (CLF) models, which aim to constrain
$\Phi(L|M)$, the luminosity (or stellar mass) function of galaxies in
haloes of mass $M$ \citep[e.g.,][]{yang03,vdBosch03a,cooray06}. The
HOD and CLF models are mainly used in combination with the halo model,
which describes the non-linear matter field in terms of halo building
blocks \citep[e.g.,][]{seljak00,cooray02}, to predict the clustering
of galaxies as function of luminosity or stellar mass.  Although
entirely analytical, and therefore fast, the downside of this method
is that it is extremely challenging to construct a model that is
accurate at the few percent level. The main issues are halo exclusion
\citep[see][]{smith_etal07,vdBosch13}, scale dependence of the halo
bias \citep[see][]{tinker05}, and the halo assembly bias mentioned
above.

These issues are trivially avoided by directly populating dark matter
haloes in numerical simulations with galaxies. This can be done using
either a parameterized model for the halo occupation statistics
\citep[e.g.,][]{Yang_etal04,moster10,behroozi10,behroozi13} or by
rank-order matching galaxy properties (i.e., luminosity or stellar
mass) to halo properties (i.e., halo mass). The latter method is known
as subhalo abundance matching (SHAM).  Although many SHAM studies use
halo mass as the halo property of choice
\citep[e.g.,][]{vale_ostriker04,vale_ostriker06,conroy07,conroy_wechsler09},
a number of studies have used alternative halo properties, such as the
maximum circular velocity, $\vmax$, or its peak value over the halo's
history, $V_{\rm peak}$
\citep[e.g.][]{tasitsiomi_etal04,conroy06,marin_etal08,reddick_etal13}

Since, at fixed halo mass, $\vmax$ is correlated with halo
concentration, and thus with halo assembly time, the resulting
occupation models have assembly bias build in
\citep[see][]{zentner_etal14}; i.e., at fixed halo mass, galaxy
properties will be correlated with halo assembly
time. \citet{lehmann_etal15} took this one step further, and explored
SHAM models in which the halo property varies between $M$ and $\vmax$
via a continuously-valued parameter. Using SDSS clustering
measurements to constrain their model, they conclude that galaxy
luminosity reveals a significant dependence on halo
concentration. Since the latter is strongly correlated with halo
formation time \citep[e.g.,][]{wechsler02,zhao_etal09,ludlow_etal13},
their conclusions argue for non-zero assembly bias in the galaxy
population.

@@

Our work is important for several reasons. Our work is a re-analysis
of the SDSS DR7 data that overcomes shortcomings of previous
analyses. In particular, we use direct population of galaxies in a
cosmological simulation so that delicate issues present in analytic
modeling, such as scale-dependent halo bias, are treated exactly.  The
simulation we use is based on the latest Planck cosmological
parameters, updating previous work. Furthermore, many important
differences between this work and the previous analysis by
\citet{zehavi_etal11} result from the fact that the Monte Carlo Markov
Chains used by Zehavi et al. were too small and had therefore not
converged (Z. Zheng \& I. Zehavi, private communication).  Therefore,
our analysis {\em supersedes previous analyses even in the case of
  standard halo occupation models}.
